---
title: "In-Class Exercise 03: Geospatial Data Science"
author: "Chong Zhi Cheng"
date: "September 2, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
  warning: false
---

### Getting Started

Note that this file will not contain all the code mentioned in class as I felt that it was more beneficial to listen and understand instead of trying to copy everything down.

Have set eval:false for this page.

-   Note that maptools is depreciated. However, this is a way to find depreciated packages.

#### Install and Launch R Packages

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```

```{r}
#| eval: false
install.packages("maptools",repos='https://packagemanager.posit.co/cran/2023-10-13')
```

The various differences between st_combine and st_unions

```{r}
#| eval: false
sg_sf <- mpsz_sf %>% st_union()

```

Note that the reason why the previous as.SpatialGridFrame.im didnt work was because maptools wasnt installed.

Hence, we had to do the workaround mentioned in Piazza

```{r}
#| eval: false
gridded_kde_childcareSG_ad <- maptools::as.SpatialGridDataFrae.im(kde_childcareSG_adaptive)
spplot(gridded_kde_childcareSG_ad)
```

Interesting data : ( also in index)

<https://acleddata.com/data-export-tool/>

themimu for more info

Importing the data from the ACLED_Myanmar csv.

Note that you should check the type of format the data is imported from. eg.

Event Date, 31 December 2021 can be interpreted as char instead of datetime.

```{r}
#| eval: false
st_as_sf(coords=c("logntiude", "latitude"),crs=4326) %>% st_transform(crs = 32647) %>%mutate(event_date=dmy(event_date))
```
