---
title: "In-class Exercise 05: Spatial Weights and Applications"
author: "Chong Zhi Cheng"
date: "September 16, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

::: callout-note
This is a note callout. I will use this for personal comments as well.
:::

### **The Data**

-   Two data sets will be used in this hands-on exercise, they are:

    -   Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.

    -   Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

## **Getting Started**

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr, GWmodel)
```

## **Getting the Data Into R Environment**

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

```{r}
hunan <- left_join(hunan, hunan2012) %>% 
  select(1, 3, 7, 15, 16, 31, 32)
```

```{r}
write_rds(hunan, "data/rds/hunan.rds")
```

```{r}
hunan <- read_rds("data/rds/hunan.rds")
```

```{r}
hunan_sp <- hunan %>% as_Spatial()
```

Adaptive Bandwidth

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "AIC",
  adaptive = TRUE,
  kernel = "bisquare",
  longlat = T
)
```

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "CV",
  adaptive = TRUE,
  kernel = "bisquare",
  longlat = T
)
```

Fixed

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "AIC",
  adaptive = FALSE,
  kernel = "bisquare",
  longlat = T
)
```

```{r}
bw_AIC <- bw.gwr(
  GDPPC ~1,
  data = hunan_sp,
  approach = "CV",
  adaptive = FALSE,
  kernel = "bisquare",
  longlat = T
)
```

```{r}
gwstat <- gwss(
  data = hunan_sp,
  vars = "GDPPC",
  bw = bw_AIC,
  adaptive = TRUE,
  kernel = "bisquare",
  longlat = T
)
```

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

```{r}
hunan_gstat <- cbind(hunan, gwstat_df)
```

```{r}
tm_shape(hunan_gstat) +
  tm_fill(
    "GDPPC_LM",
    n = 5,
    style = "quantile"
  ) +
  tm_borders(alpha = 0.5) +
  tm_layout(
    main.title = "Distribution of geographically weighted mean",
    main.title.position = "center",
    legend.text.size = 1.2,
    legend.height = 1.50,
    legend.width = 1.50,
    frame = TRUE
  )
```

## **References**
