---
title: "In-Class Exercise 04: Spatio-Temporal Point Patterns Analysis"
author: "Chong Zhi Cheng"
date: "September 9, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
  warning: false
---

### Getting Started

Note that this file will not contain all the code mentioned in class as I felt that it was more beneficial to listen and understand instead of trying to copy everything down.

A spatio-temporal point process (also called space-time or spatial-temporal point process) is a random collection of points, where each point represents the time and location of an event. Examples of events include incidence of disease, sightings or births of a species, or the occurrences of fires, earthquakes, lightning strikes, tsunamis, or volcanic eruptions.

The analysis of spatio-temporal point patterns is becoming increasingly necessary, given the rapid emergence of geographically and temporally indexed data in a wide range of fields. Several spatio-temporal point patterns analysis methods have been introduced and implemented in R in the last ten years. This chapter shows how various R packages can be combined to run a set of spatio-temporal point pattern analyses in a guided and intuitive way. A real world forest fire events in Kepulauan Bangka Belitung, Indonesia from 1st January 2023 to 31st December 2023 is used to illustrate the methods, procedures and interpretations.

### **The research questions**

The specific question we would like to answer is:

-   are the locations of forest fire in Kepulauan Bangka Belitung spatial and spatio-temporally independent?

-   if the answer is NO, where and when the observed forest fire locations tend to cluster?

### **The data**

For the purpose of this exercise, two data sets are used, they are:

-   forestfires, a csv file provides locations of forest fire detected from the Moderate Resolution Imaging Spectroradiometer (MODIS) sensor data. The data are downloaded from [Fire Information for Resource Management System](https://firms.modaps.eosdis.nasa.gov/download/). For the purpose of this exercise, only forest fires within Kepulauan Bangka Belitung will be used.

-   Kepulauan_Bangka_Belitung, an ESRI shapefile showing the sub-district (i.e. [kelurahan](https://en.wikipedia.org/wiki/Villages_of_Indonesia)) boundary of Kepulauan Bangka Belitung. The data set was downloaded from [Indonesia Geospatial](https://www.indonesia-geospasial.com/2023/05/download-shapefile-batas-administrasi.html) portal. The original data covers the whole Indonesia. For the purpose of this exercise, only sub-districts within Kepulauan Bangka Belitung are extracted.

## **Installing and Loading the R packages**

For the purpose of this study, five R packages will be used. They are:

-   rgdal for importing geospatial data in GIS file format such as shapefile into R and save them as Spatial\*DataFrame,

-   maptools for converting Spatial\* object into ppp object,

-   raster for handling raster data in R,

-   spatstat for performing Spatial Point Patterns Analysis such as kcross, Lcross, etc., and

-   tmap for producing cartographic quality thematic maps.

#### Install and Launch R Packages

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse)
```

```{r}
pacman::p_load(spatstat)
```

Importing and Preparing Study Area

Importing study area

```{r}
kbb <- st_read(dsn = 'data/rawdata', layer = "Kepulauan_Bangka_Belitung") %>%
  st_transform(crs = 32748)
```

```{r}
kbb_sf <- st_read(dsn = 'data/rawdata', layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

Converting OWIN

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

```{r}
class(kbb_owin)
```

## Importing and Preparing Forest Fire data

Next, we will import the for

```{r}
fire_sf <- read.csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 32748)


```

```{r}
fire_sf
```

```{r}
fire_sf <- fire_sf %>%
  mutate(DayOfYear = yday(acq_date)) %>%
  mutate(month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, label = TRUE, abbr = FALSE))
```

```{r}
tm_shape(kbb_sf) +   
tm_polygons() +
tm_shape(fire_sf) + 
tm_dots()
```

```{r}
tm_shape(kbb_sf) +   
tm_polygons() +   
tm_shape(fire_sf) +   
tm_dots(size=0.1) +   
tm_facets(by="Month_fac",free.coords=FALSE,drop.units=TRUE)
```

```{r}
fire_month <- fire_sf |>
  select(month_num)

fire_month
```

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin] 
summary(fire_month_owin)
```

```{r}
plot(fire_month_owin)
```

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

```{r}
tims <- c(7, 8, 9, 10, 11, 12)

par(mfcol = c(2, 3))

for(i in tims) {
  plot(st_kde, i,
       override.par = FALSE,
       fix.range = TRUE,
       main = paste("KDE at month", i))
}
```
